---
title: "R Notebook"
output: html_notebook
---

```{r}
library(tidyverse)
library(Seurat)
library(scDblFinder)
library(SingleCellExperiment)
options(Seurat.object.assay.version = "v5")
options(future.globals.maxSize = 8000 * 1024^2)
```

# Load data
```{r}
mat <- Read10X("../data/snPATHO/PCa_1/sample_filtered_feature_bc_matrix/")
pca1 <- CreateSeuratObject(counts = mat,
                           project = "PCa_1")

mat <- Read10X("../data/snPATHO/PCa_2/sample_filtered_feature_bc_matrix/")
pca2 <- CreateSeuratObject(counts = mat,
                           project = "PCa_2")

seurat <- merge(pca1, pca2)
seurat[["RNA"]] <- JoinLayers(seurat[["RNA"]])
```

# Doublet removal
Have looked at data without explicit doublet prediction and there are clear doublet populations. They're not overly problematic, but could be cleaned up

```{r}
sce <- SingleCellExperiment(list(counts = seurat[["RNA"]]$counts))
sce$sample_id <- seurat$orig.ident
sce <- scDblFinder(sce, samples = "sample_id", dbr=0.1)
```

```{r}
seurat$doublet <- sce$scDblFinder.class
```


# QC
```{r}
seurat$percent_mt <- PercentageFeatureSet(seurat, pattern="^MT-")
seurat$nCount_RNA_log10 <- log10(seurat$nCount_RNA)
```

```{r}
hist(seurat$nCount_RNA_log10, breaks=100)
hist(seurat$nFeature_RNA, breaks=100)
```

First looked at data without any filtering. Low count/feature droplets seem non-specific--perhaps debris. Have immunoglobulins and no specific markers

```{r}
seurat <- subset(seurat, nFeature_RNA > 1000 & doublet == "singlet")
```


# Processing
```{r}
seurat[["RNA"]] <- JoinLayers(seurat[["RNA"]])
seurat <- SCTransform(seurat) %>%
  RunPCA(verbose=F) %>%
  RunUMAP(dims=1:30) %>%
  FindNeighbors(dims=1:30) %>%
  FindClusters(resolution=0.2)
```

```{r}
DimPlot(seurat, label=T)
```

Have looked into cluster 4/15 and both seem to clearly be debris. No specific markers
```{r}
seurat <- subset(seurat, seurat_clusters %nin% c(4, 15))
```

```{r}
seurat <- SCTransform(seurat) %>%
  RunPCA(verbose=F) %>%
  RunUMAP(dims=1:30) %>%
  FindNeighbors(dims=1:30) %>%
  FindClusters(resolution=0.2)
```

```{r}
DimPlot(seurat, label=T)
```

# Load atlas
```{r}
ref <- readRDS("../data/snPATHO/atlas/prostate_portal_300921.RDS")
```

```{r}
common_genes <- intersect(rownames(ref), rownames(seurat))
ref <- CreateSeuratObject(counts = assay(ref, "soupx_counts"),
                          meta = as.data.frame(colData(ref)))
```

```{r}
ref <- NormalizeData(ref) %>%
  FindVariableFeatures() %>%
  ScaleData() %>%
  RunPCA(verbose=F) %>%
  RunUMAP(dims=1:30)
```


```{r}
DefaultAssay(seurat) <- "RNA"
seurat <- NormalizeData(seurat)

anchors <- FindTransferAnchors(reference = ref, 
                                query = seurat,
                                dims = 1:30, 
                                reference.reduction = "pca")
predictions <- TransferData(anchorset = anchors, 
                            refdata = ref$celltype,
                            dims = 1:30)
seurat$celltype_pred <- predictions$predicted.id
```

```{r}
DimPlot(seurat, group.by='celltype_pred', label=T)
DimPlot(seurat, label=T)
```

Add annotations
```{r}
anno <- data.frame(
  cluster = 0:14,
  celltype = c("Epithelial (Luminal)",
               "Epithelial (Luminal)",
               "Smooth muscle",
               "Fibroblast",
               "Epithelial (Basal)",
               "Endothelial",
               "Epithelial (Luminal)",
               "Epithelial (Luminal)",
               "Epithelial (Club)",
               "Pericyte",
               "Muscle (CHRNG+)",
               "Macrophage",
               "T/NK cell",
               "Epithelial (DPYS+)",
               "Mast cell")
)
seurat$celltype <- anno$celltype[match(seurat$seurat_clusters, anno$cluster)]
```

```{r}
DimPlot(seurat, group.by='celltype', label=T)
```

## Add lower-res annotation
```{r}
anno <- data.frame(
  celltype = c("Epithelial (Luminal)",
             "Epithelial (Basal)",
             "Epithelial (Club)",
             "Epithelial (DPYS+)",
             "Endothelial",
             "Fibroblast",
             "Smooth muscle",
             "Muscle (CHRNG+)",
             "Pericyte",
             "T/NK cell",
             "Macrophage",
             "Mast cell"),
  celltype_lineage = c("Epithelial", 
                       "Epithelial", 
                       "Epithelial", 
                       "Epithelial",
                       "Endothelial",
                       "Fibroblast",
                       "Muscle",
                       "Muscle",
                       "Pericyte",
                       "T/NK cell",
                       "Macrophage",
                       "Mast cell")
)

seurat$celltype_lineage <- anno$celltype_lineage[match(seurat$celltype, anno$celltype)]
```


# Save point

```{r}
saveRDS(seurat, file="../output/snPATHO_processed.rds")
```


